### 2025-03-23 paper/paper.tex
initial commit

\documentclass[12pt]{article}

% Packages
\usepackage[utf8]{inputenc}
\usepackage{xeCJK}
\usepackage{ctex}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{float}
\usepackage{cite}

% Page layout
\geometry{a4paper, margin=1in}

% Title and Author
\title{Final Paper Title}
\author{Your Name}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
This is the abstract of your paper. Provide a brief summary of the content, objectives, and conclusions of your work.
\end{abstract}

\section{Introduction}
Introduce the topic of your paper. Provide background information and state the objectives of your work.

\section{Related Work}
Discuss previous research or work related to your topic. Cite relevant sources, such as \cite{GOYAL2020220}.

\section{Methodology}
Explain the methods or approaches used in your work. Provide enough detail for reproducibility.\cite{Fan2019}

\section{Results}
Present the results of your work. Use tables, figures, and equations as needed.\cite{中值滤波}

\section{Discussion}
Interpret the results and discuss their implications. Compare with related work if applicable.

\section{Conclusion}
Summarize the main findings of your paper and suggest future work.

\appendix
\section{Appendix}
Include additional material, derivations, or data that support your paper but are not essential to the main text.

\bibliographystyle{plain}
\bibliography{ref}

\end{document}